### 2025-04-17 paper/paper.tex
Wrote more for the paper

\documentclass[conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
%Template version as of 6/27/2024

\usepackage{cite}
\usepackage{cprotect}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{outlines}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\graphicspath{ {./images/} }
\begin{document}

\title{A Lightweight, Swappable Voxel Ray-Tracer}

\makeatletter
\newcommand{\linebreakand}{%
  \end{@IEEEauthorhalign}
  \hfill\mbox{}\par
  \mbox{}\hfill\begin{@IEEEauthorhalign}
}
\makeatother

\author{\IEEEauthorblockN{Ann Gao}
\IEEEauthorblockA{\textit{Dept. Computer Science} \\
\textit{University of Central Florida}\\
Orlando, FL \\
an863834@ucf.edu}
\and
\IEEEauthorblockN{Nathan Hall}
\IEEEauthorblockA{\textit{Dept. Computer Science} \\
\textit{University of Central Florida}\\
Orlando, FL \\
na116376@ucf.edu}
\and
\IEEEauthorblockN{Jonah Henriksson}
\IEEEauthorblockA{\textit{Dept. Computer Science} \\
\textit{University of Central Florida}\\
Orlando, FL \\
jo295297@ucf.edu}
\linebreakand
\IEEEauthorblockN{Christopher Jenkins}
\IEEEauthorblockA{\textit{Dept. Computer Science} \\
\textit{University of Central Florida}\\
Orlando, FL \\
ch119158@ucf.edu}
\and
\IEEEauthorblockN{Felipe Schmidt}
\IEEEauthorblockA{\textit{Dept. Computer Science} \\
\textit{University of Central Florida}\\
Orlando, FL \\
fe805154@ucf.edu}
}

\maketitle

\begin{abstract}
Voxel rendering is a technique for visualizing 3D volumes, with applications in medical technology, terrain visualization, and video games.
Simple voxel rendering implementations rely on “meshing”, a process by which the volume is converted to a mesh for rasterization.
For many tasks, this is sufficient, but also suffers from memory constraints which prevents meshing from scaling to large scenes (since both the voxel data and heavy mesh representation has to be stored).
An alternative solution that lacks such constraints is ray-tracing; which operates on the voxel data itself.
In this project, we present a CPU voxel ray-tracer implemented in Rust, alongside two storage solutions that can be swapped: a dense array-backed storage, and a sparse octree storage.
The relative performance of these solutions are assessed via render benchmarks.
\end{abstract}

\begin{IEEEkeywords}
voxels, ray-tracing, octree.
\end{IEEEkeywords}

\section{Introduction}
The term “voxels” is derived from the term “pixels”: where a “pixel” is a “\textbf{pic}ture \textbf{el}ement” that represents part of a picture, a “voxel” is a “\textbf{vo}lume \textbf{el}ement” that represents part of a volume.
Volumetric data can range from MRI slices to point clouds to generated volumes in video games.
Voxels are a representation of this volumetric data in a grid structure, just as pixels store raster images in a grid structure.
However, while pixels and voxels are similar in structure, the way they are visualized differs vastly.
Pixels can simply be drawn to a screen surface, but voxels are points in 3D space and drawing techniques can either draw the individual points, as with point cloud rendering, or the isosurface, which is the concern of this paper.

Isosurface rendering is the ubiquitous form of voxel rendering, to the point that “voxel rendering” almost always refers to rendering the isosurface of a voxel grid.
The isosurface of a voxel grid is the surface present between voxels that surpass some threshold and those which don’t (i.e. the boundary between voxels that are “present” and those which aren’t).
To render such an isosurface, the approach varies on the underlying render technique used.
Such render techniques generally fall into two categories: rasterization and ray-tracing.

Ray-tracing is the earliest method of rendering, which works by simulating rays of light in a scene, making it capable of rendering realistic images.
However, it is computationally expensive, leading to the popularity of another method for real time applications: rasterization.
Rasterization works by projecting triangles of a polygonal mesh onto the camera’s 2D view and filling in their bounds using small programs called “shaders”.
Because of the strengths and weaknesses of both methods, voxel rendering techniques fall into two categories as well: isosurface mesh extraction for rasterization and voxel traversal for ray-tracing (there are also hybrid techniques, but they exist as an optimization of voxel traversal).

Isosurface mesh extraction, popularly known as “meshing”, is the process of converting voxels to a mesh of the isosurface, and there are many meshing algorithms, such as “greedy meshing”, “marching cubes”, “naive surface nets”, etc.
All have their own characteristics making them suitable for particular applications.
However, when rendering large scenes, meshing and rasterization typically struggle to scale in performance.
Meanwhile, ray tracing is capable of scaling in performance, due to optimizations that are relatively easy to implement with ray tracing, such as acceleration structures and LOD (level-of-detail), and the lack of a mesh generation step when updating voxels, as well as better memory scaling (storing a mesh takes up memory on top of storing the voxel volume).
Because of this difference in scaling potential, when rendering large scenes on mid to high end consumer computers, ray tracing becomes more performant than rasterization.

\subsection{Problem Statement}

\textbf{We have achieved the goal for this project: implement a CPU voxel ray-tracer in the Rust language with two storage backends: a dense array storage and a sparse voxel octree storage, and benchmark them against each other.}
The dense array represents a naive voxel storage implementation, while the sparse octree represents an acceleration structure implementation: an optimization for ray traversal via a space partitioning structure.
Using both storage solutions, we can assess the benefits of octrees for optimizing ray traversal in voxel scenes via benchmarks simulating rendering scenes at different sizes and resolutions.
See Fig.~\ref{render} for an example render produced by our voxel ray-tracer.

\begin{figure}[htbp]
\centerline{\includegraphics[width=\linewidth]{render}}
\caption{A render produced by our voxel ray-tracer.}
\label{render}
\end{figure}

\section{Methodology}

By making the ray tracer generic over storage backends, we can swap between ray tracing implementations.
To assess the performance of each implementation, the benchmarks will consist of measuring the time it takes for each implementation to render a scene for a given scene size and frame resolution.

\section{Implementation}

The program is divided into the following parts:

\begin{outline}
\1 Voxel Generator
\1 Scene Object with Swappable Storage Backends
\2 Dense Array-Backed Storage
\2 Sparse Octree-Backed Storage
\1 Framebuffer
\1 Camera Controller
\1 Ray Tracer
\1 Image Exporter
\1 Command Line Interface
\end{outline}

The clear boundaries between components was beneficial to dividing work among colleagues.

\subsection{Voxel Generation}

[Elaborate on voxel generation]

\subsection{Scene Object}

The scene object stores the scene and provides an interface for ray tracing.
Our project has two scene object implementations: a dense array-backed storage object and a sparse octree-backed storage object.
There is some overlap in the implementations of these objects, particularly the \verb|Scene| trait, \verb|Ray| struct, and \verb|IAabb| struct:

\cprotect\subsubsection{\verb|Scene| trait}

The \verb|Scene| trait has two items: a function for creating the scene object given the voxel generator and a method for tracing a ray through the scene.
This is the interface from which the ray tracer interacts with the scene.

\cprotect\subsubsection{\verb|Ray| struct}

The \verb|Ray| struct is a vector pair of the ray origin and the ray direction.
It is used to simulate light rays that traverse the scene from the camera.

\cprotect\subsubsection{\verb|IAabb| struct}

The \verb|IAabb| struct represented an axis-aligned bounding box (AABB) represented by integer vectors.
By iterating over every coordinate in an AABB, the storage backend can query the voxel generator for a voxel and store it if present.

An AABB can also be used during a ray trace, as a quick intersection test for the entire volume.
The algorithm used for fast intersection tests is from "An Efficient and Robust Ray-Box Intersection Algorithm" by Williams et al. \cite{williams}
To summarize how the algorithm works: it takes advantage of properties observed in intersection between rays and AABBs to quickly check if the ray crosses the box boundaries.

A more detailed but brief explanation is that the algorithm projects boundary line intersections to a 1D line representing distance traveled across the ray and checks the order of these intersections to verify that the ray enters all bounds before exiting any other bounds (See Fig.~\ref{fast-intersection} for an example of the calculations).

\begin{figure}[htbp]
\centerline{\includegraphics[width=\linewidth]{fast-intersection}}
\caption{Two example calculations for fast ray-box intersections using 2D coordinates.}
\label{fast-intersection}
\end{figure}

\subsection{Dense Storage}

The dense storage backend stores every possible position in the voxel volume in a single array.
It is most beneficial to use such a storage solution when most positions hold present voxels, or in other words, is densely populated, hence the name.

\begin{figure}[htbp]
\centerline{\includegraphics[width=\linewidth]{fast-traversal}}
\caption{The fast voxel traversal algorithm in action.}
\label{fast-traversal}
\end{figure}

For the dense storage backend, there exist many algorithms for ray tracing through a multidimensional array, but most are based on the “Fast Voxel Traversal” algorithm by Amanatides and Woo, which will be what we use for ray tracing in the dense storage backend \cite{amanatides}.
This algorithm works by tracking the index and position for each axis and moving the ray along the axis with the shortest distance to the next index (See Fig.~\ref{fast-traversal}).
By doing so, it ensures that no voxels are missed during iteration while not testing positions that don't intersect the ray.
When a voxel is found or the indices go out of bounds, it returns.

\subsection{Sparse Storage}

[Sparse: Octree space partitioning]

In the world of 2D graphics there is a data structure known as a quadtree which is used to divide a 2D plane into 4 quadrants one to represent top left, top right, bottom left, and bottom right.
These quadrants are stored as children of a parent node and so on and so forth.
An Octree is the 3D equivalent of this.
Within each Octree is a root that represents an area in 3D space, within this root there are 8 children which are 8 subdivisions of the parent root, each subsequent child breaks down the space until the children being stored are individual voxels.

An octree is a helpful tool in rendering 3D spaces as they are a very compact way of storing voxels without wasting memory.
They are also very quick to traverse as one can navigate to a point in a 3D space by either descending or ascending the structure to pinpoint any given location.
One of the key features is the PreOrder Traversal, this is helpful because it can help to determine line of sight for an object, or for our purposes, a ray.
Since the ray can check what objects it will or will not collide with by simply traversing the octree it greatly reduces a lot of redundant calculations that would have been performed were the voxels to be stored in a different data structure.
The key difference between Dense and Sparse storage however is that

\subsection{Framebuffer and Image Exporter}

Once we have processed all the scene information, it is time to save it to an image so we can view it.
We do this by adding information to a framebuffer, then copying the framebuffer into a PNG file.
The framebuffer is a data structure that holds all the image information of the rendered scene.
It contains a memory reference to a heap-allocated array of 4-byte integers for each pixel.
Each integer represents the RGBA values of a pixel, with the first byte holding the R value, second byte holding the G value, etc.
The array uses an atomic data type, which allows for multiple threads to access and modify the values of different pixels safely and concurrently to allow faster image generation.
We use Rust’s \verb|image| crate to then write to an image file specified by a given output path.

\subsection{Camera Controller}

[Elaborate on camera and ray instantiation]

\subsection{Ray Tracer}

[Elaborate on how ray tracer brings parts together]

\subsection{Command Line Interface}

[Elaborate on the CLI]

\section{Evaluation}

To evaluate the performance benefits of sparse voxel octrees over array-backed storage, we constructed benchmarks via \verb|criterion.rs| that measured the performance of each storage backend over scene size and frame resolution.
The benchmarks ran across 12 threads on a mid-range consumer PC with an AMD Ryzen 5 2600 6-core CPU and 32GB of RAM.

\begin{table}[htbp]
\caption{1080p Resolution Benchmarks}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Scene Size} & \textbf{Dense} & \textbf{Sparse}  & \textbf{Ratio} \\\hline 
50 & 37.557 ms & 211.86 ms & 0.1773 \\\hline
100 & 192.48 ms & 247.86 ms & 0.7766 \\\hline
250 & 811.93 ms & 122.55 ms & 6.6253 \\\hline
\end{tabular}
\label{1080p-bench}
\end{center}
\end{table}

\begin{table}[htbp]
\caption{4k Resolution Benchmarks}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Scene Size} & \textbf{Dense} & \textbf{Sparse}  & \textbf{Ratio} \\\hline 
50 & 261.11 ms & 2.5230 s & 0.1035 \\\hline
100 & 3.0203 s & 3.5526 s & 0.8502 \\\hline
250 & 10.016 s & 2.0995 s & 4.7706 \\\hline
\end{tabular}
\label{4k-bench}
\end{center}
\end{table}

\subsection{Tracing}

[Discuss analysis via tracing]

\section{Conclusion}

[Draw some conclusion that affirms that we did what we set out to do]

\begin{thebibliography}{00}
\bibitem{williams} Williams, Amy \& Barrus, Steve \& Morley, R. \& Shirley, Peter. (2005). An Efficient and Robust Ray-Box Intersection Algorithm. J. Graphics Tools. 10. 49-54. 10.1145/1198555.1198748. 
\bibitem{amanatides} Amanatides, John \& Woo, Andrew. (1987). A Fast Voxel Traversal Algorithm for Ray Tracing. Proceedings of EuroGraphics. 87.
\end{thebibliography}

\end{document}